Investigating soundscapes perception through acoustic scenes simulation.

### 2017-12-19 lafaySimulatedScenePerception.tex
more

Considering $E$ and $T$ separately, it appears that: 1) the $p@5$ obtained with $E$ is similar to the one obtained with $ET$ ; 2) the $p@5$ obtained with $T$ is systematically lower than the one obtained with $E$, by $10\%$ or $15\%$. Those results indicate that the discriminative semantic information is mainly contained in the events. These results are in line with the work done in \cite{maffiolo_caracterisation_1999} showing that humans analyze the event scenes which are composed of several sound events in a causal manner, \ie by identifying the sources, as discussed in Section~\ref{sec:simscene_sampleDataSet}.

@@

The dimension of the vectors $S$ for $E_m$ is lower than the dimension of vector $S$ for $E$, itself lower than the dimension of vector $S$ obtained when all the classes are considered ($ET$). Despite this, it appears that the $p@5$ obtained with $E_m$ is equal – or above – to the ones obtained with $E$ or $ET$, although a partial information is used in that case to describe the scenes. Moreover, in the case when the sound markers are not taken into account for the description ($E_{w/o,m}$), the performance is under those obtained with only the textures as features. This confirms that most of the semantic information allowing to differentiate between i- and ni-scenes is included in the markers.

@@

To sum up, we deduce the following points from this analysis:

\begin{enumerate}
\item unlike what we outlined with the sound levels, a semantic description of the scenes constitution in terms of presence / absence of sound sources, allows us to differentiate the two types of environments (i or ni);
\item the semantic information is mainly contained in the event sound classes;
\item only a part of the event classes, i.e. the sound markers, are useful to differentiate between the i- and ni-scenes.
\end{enumerate}

@@

Since we have extracted the typical classes of the i- and ni-scenes and verified that the distinction between them was largely dependent on the presence of these classes, we shall now investigate whether a structural description of the scenes only based on sound markers could characterize the perceived pleasantness, perhaps better than a global structural description.

@@

We evaluate the correlations between $\mathcal{A}_{scene}$ and the sound levels. In this section, the sound levels are computed by taking into account only the previously identified sound markers. We define $X_m$, the $X$ feature computed by taking into account only the sound markers. On the opposite, we define $X_b$, the $X$ feature computed by taking into account all the sound classes, except the sound markers. When the feature characterizes an i-scene (idem for a ni-scene),  only the markers identified for the i-scenes (or the ni-scenes) are considered; we named them, i- markers (or ni- markers). Results are shown in Table~\ref{tab:corrMarkers}.

@@

In conclusion, the level of ni- markers has a negative influence on the pleasantness for the ni-scenes. On the other hand, the level of i- markers does not influence the perceived pleasantness for the i-scenes.

@@

Besides, whereas $L(T)$ did not show any significant correlation for ni-scenes, a strong negative correlation is observed for $L(T)_b$ ($r = -0.73$, $p < 0.01$). This indicates that the level of non-markers texture classes does not influence the perceived pleasantness in the same way for i- and ni-scenes. Sound level seems to have a negative effect for the ni-scenes, and no significant effect for the i-scenes.

@@

We consider a last group of features, namely $L_m-L_b$, $L(E)_m-L(E)_b$ et $L(T)_m-L(T)_b$.  These features describe the difference between the markers level and those of the other sound classes, see ~Figures~\ref{fig:soundlevelMarkerDiffa} and~\ref{fig:soundlevelMarkerDiffd}. They express the saliency of the markers with respect to the sound mixture.

@@

For the i-scenes, a moderate positive correlation is observed for $L_m-L_b$ ($r=0.67$, $p<0.01$) and $L(E)_m-L(E)_b$ ($r=0.66$, $p<0.01$). for the ni-scenes no correlation is observed. Thus, for the i-scenes, it is not the absolute markers level that is important, but their relative level with regards to the other sounds composing the scene. Then, we can induce a double perceptual mechanism for the ideal environments:

@@

\begin{itemize}
\item the higher the absolute level of sounds not being i- markers, the weaker the pleasantness,
\item the higher the relative level of i- markers is, with regards to the other sounds, the higher the pleasantness.
\end{itemize}

@@

For the ni-scenes, the fact that we observe significant correlations for $L_m$ and $L(E)_m$ and no correlation for $L_m-L_b$ and $L(E)_m$ - $L(E)_b$ shows that this is indeed the absolute level that matters.

@@

\item \emph{differentiating the i- and ni-scenes}: the semantic features, and the global sound levels ($L$, $L(E)$ et $L(T)$), allow us to differentiate between i- and ni-scenes. The semantic description seems to be more powerful;
\item \emph{events or textures}: whatever the feature type, be it semantic or structural, events are mainly useful to differentiate the two types of scenes; textures bring a limited amount of information;
\item \emph{pleasantness prediction}: considering the correlation between sound levels and pleasantness, it seems that the way subjects perceived the quality of a given environment depends on its nature (i or ni). From the data gathered in those experiments, it is not conceivable to consider the same set of features to predict the pleasantness of i- and ni-scenes in the same time:

>>>>

Considering $E$ and $T$ separately, it appears that: 1) the $p@5$ obtained with $E$ is similar to the one obtained with $ET$ ; 2) the $p@5$ obtained with $T$ is always lower than the one obtained with $E$, by $10\%$ to $15\%$. Those results indicate that the semantic information that is discriminative is mostly carried by the events. Those results are in line with \cite{maffiolo_caracterisation_1999}. As discussed in Section~\ref{sec:simscene_sampleDataSet}, it seems that humans analyze the event scenes which are composed of several sound events in a causal manner, \ie by identifying the sources.

@@

The dimension of the vectors $S$ for $E_m$ is lower than the dimension of vector $S$ for $E$, itself lower than the dimension of vector $S$ obtained when all the classes are considered ($ET$). $S$ being a boolean vector, the smaller the dimension, the lower the amount of information it can carry. Despite this, it appears that the $p@5$ obtained with $E_m$ is equal – or above – to the ones obtained with $E$ or $ET$, although only a partial information is used in that case to describe the scenes. On the opposite, if the sound markers are not taken into account for the description ($E_{w/o,m}$), the performance is below the one achieved when considering only the textures as features. Thus, most of the semantic information allowing to differentiate between i- and ni-scenes is included in the markers.

@@

The outcomes of this analysis are:

\begin{enumerate}
\item unlike what we outlined with the sound levels, a semantic description of the scenes composition in terms of presence / absence of sound sources, allows us to differentiate reliably the two types of environments (i or ni);
\item the semantic information is mainly contained in the event sound classes;
\item only a part of the event classes, i.e. the sound markers, are useful to differentiate between the i- and ni-scenes.
\end{enumerate}

@@

Since we have extracted the typical classes of the i- and ni-scenes and verified that the distinction between them was largely dependent on the presence of these classes, we shall now investigate whether a description of the scenes only based on the sound pressure level of sound markers could characterize the perceived pleasantness, perhaps better than a globally computed sound level.

@@

To do so, the correlations between $\mathcal{A}_{scene}$ and the sound levels are evaluated. In this section, the sound levels are computed by taking into account only the previously identified sound markers. We define $X_m$, the $X$ feature computed by taking into account only the sound markers. On the opposite, we define $X_b$, the $X$ feature computed by taking into account all the sound classes, except the sound markers. When the feature characterizes an i-scene (idem for a ni-scene),  only the markers identified for the i-scenes (or the ni-scenes) are considered; we named them, i-markers (or ni- markers). Results are shown in Table~\ref{tab:corrMarkers}.

@@

To conclude, the level of ni- markers has a negative influence on the pleasantness for the ni-scenes. On the other hand, the level of i- markers does not influence the perceived pleasantness for the i-scenes.

@@

Alos, whereas $L(T)$ did not show any significant correlation for ni-scenes, a strong negative correlation is observed for $L(T)_b$ ($r = -0.73$, $p < 0.01$). This indicates that the level of non-markers texture classes does not influence the perceived pleasantness in the same way for i- and ni-scenes. Sound level seems to have a negative effect for the ni-scenes, and no significant effect for the i-scenes.

@@

A last group of features is now considered, namely $L_m-L_b$, $L(E)_m-L(E)_b$ and $L(T)_m-L(T)_b$.  These features describe the difference between the markers level and those of the other sound classes, see ~Figures~\ref{fig:soundlevelMarkerDiffa} and~\ref{fig:soundlevelMarkerDiffd}. They express the saliency of the markers with respect to the sound mixture.

@@

For the i-scenes, a moderate positive correlation is observed for $L_m-L_b$ ($r=0.67$, $p<0.01$) and $L(E)_m-L(E)_b$ ($r=0.66$, $p<0.01$). for the ni-scenes no correlation is observed. Thus, for the i-scenes, it is not the absolute markers level that is important, but their relative level with respect to the other sounds composing the scene. A double perceptual mechanism for the ideal environments: can thus be observed:

@@

\begin{itemize}
\item the higher the absolute level of sounds not being i-markers, the weaker the pleasantness,
\item the higher the relative level of i-markers is, with respect to the remaining sounds, the higher the pleasantness.
\end{itemize}

@@

On contrary, the fact that we observe significant correlations for $L_m$ and $L(E)_m$ and no correlation for $L_m-L_b$ and $L(E)_m$ - $L(E)_b$ for the ni-scenes, shows that this is indeed the absolute level that matters.

@@

\item \emph{differentiating the i- and ni-scenes}: the semantic features, and the global sound levels ($L$, $L(E)$ et $L(T)$), allow us to differentiate between i- and ni-scenes. The semantic description seems to be more powerful;
\item \emph{events or textures}: whatever the feature type, be it semantic or related to the sound pressure level, events are mainly useful to differentiate the two types of scenes; textures bring a limited amount of information;
\item \emph{pleasantness prediction}: considering the correlation between sound levels and pleasantness, it seems that the way subjects perceived the quality of a given environment depends on its nature (i or ni). From the data gathered in those experiments, it is not conceivable to consider the same set of features to predict the pleasantness of i- and ni-scenes in the same time:
